### 2019-03-04 Manuscript/social_closeness_dishonesty.Rmd
Plots
Adding plots into separate files

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/Documents/R/github_Said/social_closeness/Manuscript/social_closeness_dishonesty_files/svm-latex-ms.tex
title: "Social Closeness and Dishonesty"
thanks: "Replication files are available on the author's Github account..."
author:
- name: Said E. Jiménez
  affiliation: Universidad Nacional Autónoma de México
- name: Roberto Mercadillo
  affiliation: Universidad Autónoma Metropolitana
- name: Eduardo Garza-Villarreal
  affiliation: Instituto Nacional de Psiquiatría Ramón de la Fuente Muñiz
abstract: "This document provides an introduction to R Markdown, argues for its..."
keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
# spacing: double
bibliography: ~/Dropbox/master.bib
biblio-style: apsr
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.