### 2019-03-04 Manuscript/social_closeness_dishonesty.Rmd
From tiff to jpeg
.

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/Documents/R/github_Said/social_closeness/Manuscript/social_closeness_dishonesty_files/svm-latex-ms.tex
title: "Social Closeness and Dishonesty"
thanks: "Replication files are available on the author's Github account https://github.com/saidejp/social_closeness"
author:
- name: Said E. Jiménez
  affiliation: Universidad Nacional Autónoma de México
- name: Roberto Mercadillo
  affiliation: Universidad Autónoma Metropolitana
- name: Eduardo Garza-Villarreal
  affiliation: Instituto Nacional de Psiquiatría Ramón de la Fuente Muñiz
abstract: "This document provides an introduction to R Markdown, argues for its..."
keywords: "dishonesty, bayesian hierarchical modeling, mixture modeling"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
fig_caption: yes
# spacing: double
bibliography: ~/Dropbox/master.bib
biblio-style: apsr
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Introduction

bla bla...

# Method

![Rates](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g1.jpeg)

![Rates by partner](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g2.jpeg)

![Rates by partners and promise condition](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g3.jpeg)

![Promises](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g4.jpeg)

![Bayesians logistic coefficients](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g5.jpeg)

![Linear population predictor](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g6.jpeg)

![Linear population predictor by subject](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g7.jpeg)

![Population coefficient with uncertainty](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g8.jpeg)

![Model adjusted by subject](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g9.jpeg)

![Two latent thetas](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g16.jpeg)

![Subjects from two latent thetas](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g10.jpeg)

![Strategies by group](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g11.jpeg)

![Rates by group](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g12.jpeg)

![Rates by group and promise condition](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g12_b.jpeg)

![Reaction times by group](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g13.jpeg)

![Reaction times by partner](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g14.jpeg)

![Difference between groups in reaction time](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g15.jpeg)