### 2019-03-05 Manuscript/social_closeness_dishonesty.Rmd
.

---
# Template adapted from:http://svmiller.com/blog/2016/02/svm-r-markdown-manuscript/

output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/Documents/R/github_Said/social_closeness/Manuscript/social_closeness_dishonesty_files/svm-latex-ms.tex
title: "Social Closeness and Dishonesty"
thanks: "Replication files are available on the author's Github account https://github.com/saidejp/social_closeness"
author:
- name: Said E. Jiménez
  affiliation: Universidad Nacional Autónoma de México
- name: Roberto Mercadillo
  affiliation: Universidad Autónoma Metropolitana
- name: Eduardo Garza-Villarreal
  affiliation: Instituto Nacional de Psiquiatría Ramón de la Fuente Muñiz
abstract: "This document provides an introduction to R Markdown, argues for its..."
keywords: "dishonesty, bayesian hierarchical modeling, mixture modeling"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
# fig_caption: yes
# spacing: double
bibliography: ~/Dropbox/master.bib
biblio-style: apsr
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      out.width = '80%',
                      fig.align = 'center',
                      dpi = 300)

library(pacman)
p_load(magrittr, dplyr, forcats, tidyr, modelr, tidybayes,
       ggstance, ggridges, rstan, brms, tidyverse, bayesplot,
       rjags)

```

# Introduction

bla bla...

# Method

![Difference between groups in reaction time](/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g1.jpeg)

```{r g1, fig.cap='Difference between groups in reaction time'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g1.jpeg")

```

```{r g2, fig.cap='Rates by partner'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g2.jpeg")

```

```{r g3, fig.cap='Rates by partners and promise condition'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g3.jpeg")

```

```{r g4, fig.cap='Promises'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g4.jpeg")

```

```{r g5, fig.cap='Bayesians logistic coefficients'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g5.jpeg")

```

```{r g6, fig.cap='Linear population predictor'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g6.jpeg")

```

```{r g7, fig.cap='Linear population predictor by subject'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g7.jpeg")

```

```{r g8, fig.cap='Population coefficient with uncertainty'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g8.jpeg")

```

```{r g9, fig.cap='Model adjusted by subject'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g9.jpeg")

```

```{r g16, fig.cap='Two latent thetas'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g16.jpeg")

```

```{r g10, fig.cap='Subjects from two latent thetas'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g10.jpeg")

```

```{r g11, fig.cap='Strategies by group'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g11.jpeg")

```

```{r g12, fig.cap='Rates by group'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g12.jpeg")

```

```{r g12_b, fig.cap='Rates by group and promise condition'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g12_b.jpeg")

```

```{r g13, fig.cap='Reaction times by group'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g13.jpeg")

```

```{r g14, fig.cap='Reaction times by partner'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g14.jpeg")

```

```{r g15, fig.cap='Difference between groups in reaction time'}

knitr::include_graphics("/Users/saidjimenez/Documents/R/github_Said/social_closeness/Manuscript/figures/g15.jpeg")

```